---
title: "markdown"
author: "Ravi Prakash"
output: html_document
---
# R-Markdown for Binary Search

<br />

### Displaying the Input Array

**The following _display_ chunk displays the array i.e. used as input for the Binary search implementation** <br />
_(We just taking the array as input, as not performing any other task, in order to isplay the same as per the requirement of the question)_
```{r display}
len <- as.numeric(readline(prompt = "Enter length of the array: "))

arr <- NULL
for(i in 1:len) {
  num <- as.numeric(readline(prompt = "Enter an element: "))
  arr <- c(arr, num)
}

print(arr)
```

### Function for Binary Search

The following function takes 2 parameters viz `list` (i.e. the array) and `input` (i.e. the element to be searched) and performs the Binary Search <br />
It does **NOT** displays any output, rather the function return `TRUE` if the searched element **exists** and `FALSE`, otherwise.
```{r bsearch}
bsearch <- function(list, input){
  
  beg = 1
  end = as.numeric(length(list))
  
  while(beg <= end) {
    mid = as.integer((beg+end)/2)
    
    if(list[[mid]] == input) {
      return (TRUE)
    }
    else if(input < list[[mid]]) {
      end <- mid-1
    } else {
      beg <- mid+1
    }
    
  }
  
  return (FALSE)
}
```

### Searching an Element

The following code-chunk calls the above defined `bsearch()` function in order to search an element (8 here) in the provided array. <br />

If the function returns `TRUE` this block dosplays a message **Found!!**, else print **Not Found!!** <br />

_(We can also take the element i.e. to be searched as user-input using function `readline()`)_
```{r output}
res <- bsearch(arr, 8)

if (res) {
  print("Found!!") 
} else {
  print("Not Found!!")
}
```
