---
title: "Titanic Disaster Servivers' Analysis"
output: html_notebook
---

This is an R Notebook for classification of servivers of [Titanic Disaster](http://rmarkdown.rstudio.com), based on the different aspects like age, sex and pessanger class etc.. The training and test dataset are taken from Kaggle.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

### **Loading Datasets**
```{r}
train <- read.csv("CSV/train.csv")
test <- read.csv("CSV/test.csv")
```

### **Add column *servived* to test dataset**
```{r}
test.survived <- data.frame(test[,1], Survived = rep("None", nrow(test)), test[,2:11])
names(test.survived)[1] <- names(train[1])
```

```{r}
# joining data sets
data.combined <- rbind(train, test.survived)
```

### **working with some R datatypes**
```{r}
str(data.combined)
```

```{r}
data.combined$Survived <- as.factor(data.combined$Survived)
data.combined$Pclass <- as.factor(data.combined$Pclass)
```

```{r}
# gross servival rate for titanic disaster ---> a table
table(data.combined$Survived)
```

```{r}
# distribution accross classes of passengers
table(data.combined$Pclass)
```

### **Loading library for Visualization**
```{r}
library(ggplot2)
```

```{r}
# pay more, get higher place, be closer to life bags   --->
        #   checking no of people in 3 classes and no of survivers

#train$Survived <- as.factor(train$Survived)
#train$Pclass <- as.factor(train$Pclass)
table(train$Pclass)
table(train$Survived)
```

## **Plotting**
```{r}
# devision of survivers in each class
ggplot(train, aes(x = Pclass, fill = factor(Survived))) +
  geom_histogram(binwidth = 0.5) +
  xlab("Pclass") +
  ylab("Total Count") +
  labs(fill = "Survived")
```

```{r}
# separation of classes based of survival
ggplot(train, aes(x = Survived, fill = factor(Pclass))) +
  geom_histogram(binwidth = 0.5) +
  xlab("Survived") +
  ylab("Total Count") +
  labs(fill = "Pclass")
```

### **Data manipulation for Characters**
```{r}
# dealing with names in training dataset
head(as.character(train$Name))
```

```{r}
length(as.character(data.combined$Name))            # 1309
length(unique(as.character(data.combined$Name)))    # 1307    ----->  unique
# here we have 2 duplicate names
```

```{r}
# extract (2) duplicate names
dup.names <- as.character(data.combined[which(duplicated(as.character(data.combined$Name))), "Name"])
#     ----->  unique vs duplicate
```

```{r}
# look all records having duplicate names
data.combined[which(data.combined$Name %in% dup.names),]
```

```{r}
# work on string data
library(stringr)
```

```{r}
# devision based on sex & marital status (title)
misses <- data.combined[which(str_detect(data.combined$Name, "Miss.")),]
mrses <- data.combined[which(str_detect(data.combined$Name, "Mrs.")),]
masters <- data.combined[which(str_detect(data.combined$Name, "Master.")),]
misters <- data.combined[which(str_detect(data.combined$Name, "Mr.")),]

head(misses)
```

```{r}
females <- data.combined[which(data.combined$Sex == "female"),]
males <- data.combined[which(data.combined$Sex == "male"),]

head(females)
head(males)
```

### **Relation b/w title and Survival as well as Pclass**
```{r}
# extract all the possible titles
get.titles <- function(name) {
    
  name = as.character(name)
    
  # can use switch too
  if(length(grep("Miss.", name)) > 0) {
    return("Miss.")
  } else if(length(grep("Mrs.", name)) > 0) {
    return("Mrs.")
  } else if(length(grep("Master.", name)) > 0) {
    return("Master.")
  } else if(length(grep("Mr.", name)) > 0) {
    return("Mr.")
  } else {
    return("Other")
  }
}
```

```{r}
# creating a vector for all titles to append as a new column
all.titles <- NULL
for (i in 1:nrow(data.combined)) {
  all.titles <- c(all.titles, get.titles(data.combined[i, "Name"]))
}
```

```{r}
data.combined$Titles <- as.factor(all.titles) # appending new 'title' column
  nlevels(data.combined$Titles)                 # no of all titles
```

```{r}
# visualization       --->      servival per class for every title
  
#  Differentiate b/w histogram and bargraph, after running the plot, below
ggplot(data.combined[1:819,], aes(x = Titles, fill = Survived)) +
  geom_bar(width = 0.5) +
  facet_wrap(~Pclass) +
  ggtitle("Pclass") +
  xlab("Titles") +
  ylab("Total") +
  labs(fill = "Survived")
```

#### *End, for now!!*
